Manuscript_analyses

### 2022-05-12 Manuscript_analyses.Rmd
started a new rmd for manuscript. some things not working yet.

---
title: "Greenhouse gas emissions model - making scenarios"
author: "Cory Whitney and Eike Luedeling"
bibliography:
 - references/packages.bib
 - references/papers.bib
output: html_document
---

We're defining the function as in the ['Greenhouse gas emissions model'](https://github.com/ToddRosenstock/ghg-uncertainty) document. Please look there for more information about the original model design.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=FALSE, include=FALSE}
#install and load packages
library(DiagrammeR)
# devtools::install_github("eikeluedeling/decisionSupport")
library(decisionSupport)
library(ggplot2)
library(kableExtra)
library(magrittr)
library(plyr)
library(rmarkdown)
# devtools::install_github("CWWhitney/uncertainty")
library(uncertainty)
library(chillR)
library(raster)

@@

Here we load the library and the data. The data contains ranges of possible values for the simulation. Each variable is represented with a confidence interval range and a specification of the distribution shape.

```{r load_data, include=FALSE}

@@

```{r ghg_emissions_function, include=FALSE}
# Generate the GHG emissions function 
source("GHG_emissions_function.R", local = knitr::knit_global())
```

## Running the model with scenarios

The normal `mcSimulation` function takes random draws for each of the input variables from the distributions specified in the input table. The distributions that are used are the same for every run of the Monte Carlo simulation. Sometimes it's desirable to specify particular scenarios *a priori*, e.g. to simulate outcomes for particular farm types, climate scenarios etc. To do this, we would normally have to define separate input tables and run separate simulations. This is quite inconvenient. Here, we usep a function that can do this job automatically.

@@

The .csv file for the `scenarios` option specifies variables that should be modified for each of the scenarios, as well as the number of model runs for for it. Each column beyond the first two columns in the file contain information for a particular scenario. This is restricted to the parameters that should be *changed* for this scenario. Here's an example `scenarios` file:

@@

@@

Here's the resulting file:

```{r}
knitr::kable(read.table("data/farm_scenarios.csv", sep = ",", fileEncoding = "UTF-8-BOM",header=TRUE)[,1:5])

@@

Here we run the model with the `scenario_mc` function. All variables that aren't specified in the file called in the `scenarios` option remain at their original values, given in the usual input table, called with `base_estimate`.

@@

Here we plot the distributions of the various outputs from the model.

```{r plot_distributions, fig.width=10, fig.height=10}
# enteric_CH4=enteric_CH4,
enteric_CH4_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation_scenarios,
                   vars = "enteric_CH4",
                   method = "boxplot_density",
                   old_names = "enteric_CH4",
                   new_names = "Total Methane Emissions from Livestock by Enteric Fermentation") + 
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank())

@@

enteric_CH4_dist_plot + milk_yield_dist_plot + 
  enteric_CH4_CO2eq_dist_plot + mm_CH4_CO2eq_dist_plot + 
  mm_N2O_CO2eq_dist_plot + feed_CO2_dist_plot + 
  on_farm_dist_plot + 
plot_layout(guides = "collect", design = layout) 
```

# Uncertainty for per-head data

@@

```{r uncertainty_package}
uncertainty::varscatter(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
```
```{r varkernel}
uncertainty::varkernel(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
```

@@

```{r varkernelslicerange}
uncertainty::varkernelslicerange(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
                                 out_var = GHG_simulation_scenarios$y$pc_on_farm, 
                                 min_in_var = round(min(GHG_simulation_scenarios$y$pc_milk_yield)),
                                 max_in_var = round(max(GHG_simulation_scenarios$y$pc_milk_yield)), 
                                 xlab_vars = "Total emissions given milk yield range")
```

# Probability of an increase in milk yield being associated with an increase in GHG emission intensity

GHG_simulation_scenarios$y[,"emission_intensity"]<-GHG_simulation_scenarios$y$pc_on_farm/GHG_simulation_scenarios$y$pc_milk_yield

GHG_sim_scen_y<-GHG_simulation_scenarios$y

GHG_sim_scen_y<-GHG_sim_scen_y[which(!is.infinite(GHG_sim_scen_y$emission_intensity)),]

uncertainty::varkernelslicerange(in_var = GHG_sim_scen_y$pc_milk_yield,
                                 out_var = GHG_sim_scen_y$emission_intensity, 
                                 min_in_var = round(min(GHG_sim_scen_y$pc_milk_yield)),
                                 max_in_var = round(max(GHG_sim_scen_y$pc_milk_yield)), 
                                 xlab_vars = "Emission intensity (XX/L) given milk yield range")

in_var = GHG_sim_scen_y$pc_milk_yield
out_var = GHG_sim_scen_y$emission_intensity

milk_yields_to_sample<-seq(100,10000,100)
milk_yields_to_sample<-milk_yields_to_sample[which(milk_yields_to_sample<max(in_var))]
slices<-list()

# slice through the distribution at every 100 L milk yield per cow

for(i in 1:length(milk_yields_to_sample))
{slices[[i]]<-
  varslice_resample(in_var,
                      out_var,
                      expectedin_var = milk_yields_to_sample[i],
                      n_samples = 100000,
  out_var_sampling = (max(out_var) - min(out_var))/10000)$resample
 }

decreased_emissions_probability <-
  data.frame(Base_yield = NA,new_yield=NA,probability=NA)

for (i in 1:(length(milk_yields_to_sample) - 1))
{
  for (j in (i + 1):length(milk_yields_to_sample))
    {decreased_emissions_probability[nrow(decreased_emissions_probability)+1,"Base_yield"]<-milk_yields_to_sample[i]
     decreased_emissions_probability$new_yield[nrow(decreased_emissions_probability)]<-milk_yields_to_sample[j]
     decreased_emissions_probability$probability[nrow(decreased_emissions_probability)] <-
      sum(slices[[i]] > slices[[j]]) / length(slices[[i]])

}
}

decreased_emissions_probability<-decreased_emissions_probability[which(!is.na(decreased_emissions_probability$Base_yield)),]

decreased_emissions_probability[,"yield_increase"]<-decreased_emissions_probability$new_yield-decreased_emissions_probability$Base_yield

ggplot(data = decreased_emissions_probability, aes(Base_yield, yield_increase, z =
                                                      probability)) + geom_contour_filled() +
  ggtitle("Probability that emission intensity reductions occurred") + 
  xlab("Baseline milk yield per head and year (kg)") +
  ylab("Milk yield increase per head and year (kg)") +
  stat_contour(breaks = c(0.9),col="black")

@@

# Interventions

To clearly characterize the impact of interventions, we should compare their impact with a counterfactual, i.e. with the same farms without the intervention. While this is usually impossible in practice, it is easy with a model. We just have to run the simulation again with identical inputs except for those that are affected by the intervention.

We already simulated emissions for a large population of farms, and we can now apply our interventions to this population by modifying certain parameters of the input tables. What we don't want now, however, is to randomize all values again, since this would add additional noise to the results, which would mask the signal caused by the interventions.

We now first need a function that can apply our greenhouse gas emissions model to a pre-defined set of input values. We can find syntax that can achieve this in the mcSimulation function. Here it is (with slight modifications):

```{r}
run_model<-function(x_data,model)
  {
  varnames<-colnames(x_data)[which(!colnames(x_data)=="Scenario")]
  model_function_ <- function(x) {

    e <- as.list(sapply(X = varnames, FUN = function(i) as.numeric(unlist(x[i]))))
    eval(expr = body(model), envir = e)
  }
  y <- do.call(what = rbind, args = lapply(X = apply(X = x_data, 
                                                     MARGIN = 1, FUN = model_function_), FUN = unlist))
  returnObject <- list(y = data.frame(y), x = data.frame(x_data))
  returnObject$call <- match.call()
  class(returnObject) <- cbind("mcSimulation", class(returnObject))
  return(returnObject)
}

The following code can precisely reproduce the results of the previous analysis:

```{r, eval=FALSE}
res<-run_model(GHG_simulation_scenarios$x,ghg_emissions)
```

To simulate interventions, we can now feed this function with modified x_data data.frames.

## Breed intervention

Introduction of higher-performance breeds may reduce greenhouse gas emissions. We simulate an intervention that replaces animals of low-performance breeds with animals of high-performance breeds.

To retain the farm-specific performance levels obtained from the farm survey, we simulate these impacts in a relative manner. For this, we first need to determine the performance and characteristics of each of the breeds. We'll base this on the farm survey data (`Kenyaherds` dataset from above).

```{r}
cows<-Kenyaherds[which(Kenyaherds$animal_type %in% 4:6),]

ggplot(cows, aes(factor(breed),milk_yield)) +
  geom_violin(draw_quantiles=c(0.5),fill="light green") +
  xlab("Breed") + ylab("Milk yield (kg per day)") + theme_bw()

milk_summary<-cbind(aggregate(cows$milk_yield, by=list(cows$breed), FUN=median),table(cows$breed))
milk_summary<-milk_summary[,c(1,4,2)]
colnames(milk_summary)<-c("Breed","n","Median_milk_yield")
milk_summary

This evaluation shows that the difference in median milk yields is relatively small, but that the milk yield distributions show a lot of variation that is apparently not explained by the breed. It is also apparent that the number of animals these estimates are based on varies considerably. Two breeds make up `r round(sum(milk_summary$n[1:2])/sum(milk_summary$n)*100,1)` % of the entire population, with others having very few animals, which means that the medians may be somewhat unreliable (especially for the single animal of breed 7).

Let's define the intervention as follows:

All animals are replaced by high-performance animals of breed 1, which increases milk yields proportionally to the median milk yield differences. This means that milk yields are multiplied by the following factors:

```{r}
milk_summary[,"Intervention_factor"]<-sapply(milk_summary$Median_milk_yield, function(x) min(x/milk_summary$Median_milk_yield[1],1))

milk_summary             
```

A change of the breed also causes changes in weight, mature weight and weight gain.

```{r}
live_weights<-aggregate(Kenyaherds$live_weight_LW, by=list(Kenyaherds$animal_type,Kenyaherds$breed), FUN=median)
colnames(live_weights)<-c("animal_type","breed","median_live_weight")
live_weights[,"Intervention_factor"]<-NA
for(i in 1:nrow(live_weights))
  live_weights$Intervention_factor[i]<-live_weights$median_live_weight[i]/
     live_weights$median_live_weight[which(live_weights$animal_type==live_weights$animal_type[i] &                                                                                                            live_weights$breed==1)]

# mature weight should probably be changed, and the code below does this. However, the mature weight is currently the same for all the breeds, so this accomplished nothing.

mature_weights<-aggregate(Kenyaherds$mature_weight_MW, by=list(Kenyaherds$animal_type,Kenyaherds$breed), FUN=median)
colnames(mature_weights)<-c("animal_type","breed","median_mature_weight")
mature_weights[,"Intervention_factor"]<-NA
for(i in 1:nrow(mature_weights))
  mature_weights$Intervention_factor[i]<-mature_weights$median_mature_weight[i]/
     mature_weights$median_mature_weight[which(mature_weights$animal_type==mature_weights$animal_type[i] &                                                                                                            mature_weights$breed==1)]

# same for the weight gain - doesn't depend on the breed at the moment, so all the factors are 1

weight_gains<-aggregate(Kenyaherds$weight_gain_WG, by=list(Kenyaherds$animal_type,Kenyaherds$breed), FUN=median)
colnames(weight_gains)<-c("animal_type","breed","median_weight_gain")
weight_gains[,"Intervention_factor"]<-NA
for(i in 1:nrow(weight_gains))
  weight_gains$Intervention_factor[i]<-weight_gains$median_weight_gain[i]/
     weight_gains$median_weight_gain[which(weight_gains$animal_type==weight_gains$animal_type[i] &                                                                                                            weight_gains$breed==1)]
weight_gains$Intervention_factor[which(weight_gains$median_weight_gain==0)]<-1

Now we can modify the input table accordingly.

go through all the farms. Determine the 'improvement' coefficients (based on herd structure)

```{r}
farms<-unique(Kenyaherds$QQNO)
breed_intervention<-data.frame(Farm=farms, milk_yield_4=1, milk_yield_5=1, milk_yield_6=1)

Kenyaherds_interventions<-Kenyaherds[,c("QQNO",
                                        "system",
                                        "animal_type",
                                        "breed",
                                        "live_weight_LW",
                                        "mature_weight_MW",
                                        "weight_gain_WG",
                                        "milk_yield",
                                        "dig_energy",
                                        "CP.")]

Kenyaherds_interventions[,"milk_change"]<-0
cows<-which(Kenyaherds_interventions$animal_type %in% 4:6)
Kenyaherds_interventions$milk_change[cows]<- sapply(cows,function(x) 1/milk_summary$Intervention_factor[milk_summary$Breed==Kenyaherds_interventions$breed[x]]  )

milk_change<-aggregate(Kenyaherds_interventions$milk_change[cows], by=list(Kenyaherds_interventions$QQNO[cows]), FUN=mean)

breed_intervention$milk_yield_4[which(breed_intervention$Farm %in% milk_change[,1])]<-milk_change[,2]
breed_intervention$milk_yield_5<-breed_intervention$milk_yield_4
breed_intervention$milk_yield_6<-breed_intervention$milk_yield_4

for(i in 1:nrow(Kenyaherds_interventions))
  {Kenyaherds_interventions[i,"live_weight_change"]<-
    1/live_weights$Intervention_factor[which(live_weights$animal_type==Kenyaherds_interventions$animal_type[i]&
                                             live_weights$breed==Kenyaherds_interventions$breed[i])]
#   Kenyaherds_interventions[i,"mature_weight_change"]<-
#      1/mature_weights$Intervention_factor[which(mature_weights$animal_type==Kenyaherds_interventions$animal_type[i]&
#                                                 mature_weights$breed==Kenyaherds_interventions$breed[i])]
   Kenyaherds_interventions[i,"weight_gain_change"]<-
     1/weight_gains$Intervention_factor[which(weight_gains$animal_type==Kenyaherds_interventions$animal_type[i]&
                                              weight_gains$breed==Kenyaherds_interventions$breed[i])]
}

live_weight_agg<-aggregate(Kenyaherds_interventions$live_weight_change,
                           by=list(Kenyaherds_interventions$QQNO,Kenyaherds_interventions$animal_type), FUN=mean)
#mature_weight_agg<-aggregate(Kenyaherds_interventions$mature_weight_change,
#by=list(Kenyaherds_interventions$QQNO,Kenyaherds_interventions$animal_type), FUN=mean)
weight_gain_agg<-aggregate(Kenyaherds_interventions$weight_gain_change,
                           by=list(Kenyaherds_interventions$QQNO,Kenyaherds_interventions$animal_type), FUN=mean)

colnames(live_weight_agg)<-c("farm","animal_type","conv_factor")
#colnames(mature_weight_agg)<-c("farm","animal_type","conv_factor")
colnames(weight_gain_agg)<-c("farm","animal_type","conv_factor")

for(i in 1:nrow(live_weight_agg))
  breed_intervention[which(breed_intervention$Farm==live_weight_agg$farm[i]),
                     paste0("live_weight_LW_",live_weight_agg$animal_type[i])]<-
    live_weight_agg$conv_factor[i]

#for(i in 1:nrow(mature_weight_agg))
#  breed_intervention[which(breed_intervention$Farm==mature_weight_agg$farm[i]),
#                     paste0("mature_weight_LW_",mature_weight_agg$animal_type[i])]<-
#    mature_weight_agg$conv_factor[i]

for(i in 1:nrow(weight_gain_agg))
  breed_intervention[which(breed_intervention$Farm==weight_gain_agg$farm[i]),
                     paste0("weight_gain_WG_",weight_gain_agg$animal_type[i])]<-
    weight_gain_agg$conv_factor[i]

breed_intervention[is.na(breed_intervention)]<-1

go through the x file
for each row take a random draw based on adoption rate (does farm adopt or not), then (possibly) apply factors

```{r, eval=FALSE}
x_breed_intervention <- GHG_simulation_scenarios$x

adoption_rate <- 0.6

for (i in 1:nrow(x_breed_intervention))
{
  if (rbinom(1, 1, adoption_rate)) # if farmers adopt, make changes in x file
    x_breed_intervention[i, colnames(breed_intervention)[2:ncol(breed_intervention)]] <-
      x_breed_intervention[i, colnames(breed_intervention)[2:ncol(breed_intervention)]] *
      breed_intervention[which(paste0("Farm_", breed_intervention$Farm) ==
                                 x_breed_intervention$Scenario[i]),
                         2:ncol(breed_intervention)]
}
# run model for modified x file
breed_res <- run_model(x_breed_intervention, ghg_emissions)

```{r, include=FALSE, eval=FALSE}
breed_res$call<-"multi-scenario call"
save_temperature_scenarios(breed_res,"results","Breed_intervention")
```

```{r, include=FALSE}
adoption_rate <- 0.6
breed_res<-load_temperature_scenarios("results","Breed_intervention")
class(breed_res)<-"mcSimulation"
```

### Evaluate impact of breed intervention

Assumed adoption rate, based on all farms (even the ones that already have high-performance farms and therefore don't get an 'upgrade'): `r adoption_rate*100` %

```{r, fig.show="hold", out.width="50%", fig.cap="Breed intervention (left) vs. baseline (right)", message=FALSE}

uncertainty::varscatter(in_var = breed_res$y$pc_milk_yield,
                        out_var = breed_res$y$pc_on_farm,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))

uncertainty::varscatter(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))

uncertainty::varkernel(in_var = breed_res$y$pc_milk_yield,
                        out_var = breed_res$y$pc_on_farm,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))

uncertainty::varkernel(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))

# uncertainty::varkernelslicerange(in_var = breed_res$y$pc_milk_yield,
#                                  out_var = breed_res$y$pc_on_farm, 
#                                  min_in_var = round(min(breed_res$y$pc_milk_yield)),
#                                  max_in_var = round(max(breed_res$y$pc_milk_yield)), 
#                                  xlab_vars = "Total emissions given milk yield range")
# 
# uncertainty::varkernelslicerange(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
#                         out_var = GHG_simulation_scenarios$y$pc_on_farm,
#                                  min_in_var = round(min(GHG_simulation_scenarios$y$pc_milk_yield)),
#                                  max_in_var = round(max(GHG_simulation_scenarios$y$pc_milk_yield)), 
#                                  xlab_vars = "Total emissions given milk yield range")

```{r}
breed_change_in_CO2_eq_per_animal<-breed_res$y$pc_on_farm-
  GHG_simulation_scenarios$y$pc_on_farm

breed_change_in_milk_yield_per_animal<-breed_res$y$pc_milk_yield-
  GHG_simulation_scenarios$y$pc_milk_yield
```

## Feed

## Herd structure

### Retiring unproductive bulls

To simulate this scenario, we simply have to remove all animals of type 1 from the herds.

```{r , warning=FALSE, eval=FALSE}
x_nobull_intervention <- GHG_simulation_scenarios$x

nobull_adoption_rate <- 1.0

for (i in 1:nrow(x_nobull_intervention))
{
  if (rbinom(1, 1, nobull_adoption_rate)) # if farmers adopt, make changes in x file
    x_nobull_intervention$N_animal_type_1[i]<-0
}
# run model for modified x file
nobull_res <- run_model(x_nobull_intervention, ghg_emissions)

```{r, include=FALSE, eval=FALSE}
nobull_res$call<-"multi-scenario call"
save_temperature_scenarios(nobull_res,"results","noBull_intervention")
```

```{r, include=FALSE}
nobull_res<-load_temperature_scenarios("results","noBull_intervention")
class(nobull_res)<-"mcSimulation"
```

#### Evaluate impact of "Retiring unproductive bulls" intervention

We're assuming that all bulls >36 months old have been removed.

```{r, fig.show="hold", out.width="50%", fig.cap="No bulls intervention (left) vs. baseline (right)", message=FALSE}

uncertainty::varscatter(in_var = nobull_res$y$pc_milk_yield,
                        out_var = nobull_res$y$pc_on_farm,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))

uncertainty::varscatter(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))

uncertainty::varkernel(in_var = nobull_res$y$pc_milk_yield,
                        out_var = nobull_res$y$pc_on_farm,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))

uncertainty::varkernel(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))

# uncertainty::varkernelslicerange(in_var = nobull_res$y$pc_milk_yield,
#                                  out_var = nobull_res$y$pc_on_farm, 
#                                  min_in_var = round(min(nobull_res$y$pc_milk_yield,na.rm=TRUE)),
#                                  max_in_var = round(max(nobull_res$y$pc_milk_yield,na.rm=TRUE)), 
#                                  xlab_vars = "Total emissions given milk yield range")
# 
# uncertainty::varkernelslicerange(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
#                         out_var = GHG_simulation_scenarios$y$pc_on_farm,
#                                  min_in_var = round(min(GHG_simulation_scenarios$y$pc_milk_yield)),
#                                  max_in_var = round(max(GHG_simulation_scenarios$y$pc_milk_yield)), 
#                                  xlab_vars = "Total emissions given milk yield range")

### Reducing unproductive replacement males

The idea is to raise the off-take rate of replacement males by 23% (not sure why this number). We'll implement this by eliminating replacement males on all adopting farms, with a 23% adoption rate. The result should be quite similar.

```{r , warning=FALSE, eval=FALSE}
x_lessReplace_intervention <- GHG_simulation_scenarios$x

lessReplace_adoption_rate <- 0.23

for (i in 1:nrow(x_lessReplace_intervention))
{
  if (rbinom(1, 1, lessReplace_adoption_rate)) # if farmers adopt, make changes in x file
    x_lessReplace_intervention$N_animal_type_3[i]<-0
}
# run model for modified x file
lessReplace_res <- run_model(x_lessReplace_intervention, ghg_emissions)

```{r, include=FALSE, eval=FALSE}
lessReplace_res$call<-"multi-scenario call"
save_temperature_scenarios(lessReplace_res,"results","lessReplace_intervention")
```

```{r, include=FALSE}
lessReplace_res<-load_temperature_scenarios("results","lessReplace_intervention")
class(lessReplace_res)<-"mcSimulation"
```

#### Evaluate impact of "Retiring unproductive bulls" intervention

We're assuming that all bulls >36 months old have been removed.

```{r, fig.show="hold", out.width="50%", fig.cap="Fewer replacement males intervention (left) vs. baseline (right)", message=FALSE}

uncertainty::varscatter(in_var = lessReplace_res$y$pc_milk_yield,
                        out_var = lessReplace_res$y$pc_on_farm,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))

uncertainty::varscatter(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))

>>>>

---
title: "Manuscript_analyses"
author: "Eike Luedeling, Cory Whitney, Todd Rosenstock"
date: "5/12/2022"
output: html_document
---

# Accounting for uncertainty in simulation greenhouse gas emissions from dairy cattle in Kenya

## Set up

Install and load required packages

```{r setup, warning=FALSE, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

#install and load packages
# devtools::install_github("CWWhitney/uncertainty")
library(DiagrammeR)
library(decisionSupport)
library(ggplot2)
library(kableExtra)
library(magrittr)
library(plyr)
library(rmarkdown)
library(uncertainty)
library(chillR)
library(raster)

@@

```{r, warning=FALSE, eval=FALSE}
devtools::install_github("CWWhitney/uncertainty")
library(DiagrammeR)
library(decisionSupport)
library(ggplot2)
library(kableExtra)
library(magrittr)
library(plyr)
library(rmarkdown)
library(uncertainty)
library(chillR)
library(raster)
```

# Load emission calculation function

The calculations are based on the IPCC xxx equation (describe and cite here)

```{r ghg_emissions_function}
# Generate the GHG emissions function 
source("GHG_emissions_function.R", local = knitr::knit_global())
```

This equation takes a wide range of inputs. Many of these are usually not known precisely, yet the classic IPCC equation does not account for potential errors and uncertainties. Here, we express this uncertainty by defining distributions for each variable. The following table shows all the variables, as well as examples of distributions that may describe their quantities. Distributions are defined according to the distribution shape, as well as upper and lower bounds of 90% confidence intervals.

```{r load_data}

@@

## Monte Carlo simulation

With the model equation and the input table defined above, we can run a Monte Carlo simulation that returns a probability distribution for the greenhouse gas emissions resulting from the particular herd that is specified in the input table.

```{r, warning=FALSE,eval=FALSE}
GHG_emissions<-mcSimulation(estimate=estimate_read_csv(input_table),
                            model_function=ghg_emissions,
                            numberOfModelRuns=1000,
                            functionSyntax="plainNames")
```

## Running a Monte Carlo simulation with scenarios

The `mcSimulation` function takes random draws for each of the input variables from the distributions specified in the input table. The distributions that are used are the same for every run of the Monte Carlo simulation. This may be desirable in homogeneous populations where all combinations of random values are plausible, but it often misses the mark where populations are composed of particular member types with specific characteristics. To account for populations that are heterogeneous, which is the case for livestock-keeping households in Kenya, we use the `scenario_mc` function, which allows specifying particular subgroups. Defining such scenarios can be desirable when it's necessary to simulate outcomes for particular farm types, climate scenarios etc. To do this, we would normally have to define separate input tables and run separate simulations. This is quite inconvenient. Here, we use a function that can do this job automatically.

@@

The .csv file for the `scenarios` option specifies variables that should be modified for each of the scenarios, as well as the number of model runs for for it. Each column beyond the first two columns in the file contain information for a particular scenario. This is restricted to the parameters that should be *changed* for this scenario.

Here's an example `scenarios` file:

@@

We define subgroups based on the populations of household that was described by a household survey in Kenya. Each of the XXX encountered in this survey is converted into one scenario, for which Monte Carlo simulation is then used to produce multiple replicates.

## Household-based scenarios

The information for the household is extracted from a spreadsheet that contains all the survey responses. Extracting all relevant information is a slightly complex task, which is accomplished by the following code.

@@

Here's the resulting file (only the first 3 farms are shown, and only the first 20 rows):

```{r}
knitr::kable(read.table("data/farm_scenarios.csv", sep = ",", fileEncoding = "UTF-8-BOM",header=TRUE)[1:20,1:5])

@@

The first few rows in this table adjust the herd composition parameters, so that the scenario herd corresponds to the herd of the specific household. Further parameters that are adjusted include the milk yield per cow and various other factors.

Now we can run the Monte Carlo simulation with these scenarios.

@@

Can can now plot distributions of the various outputs from the model, using the `plot_distributions` function.

```{r plot_distributions, fig.width=10, fig.height=10}
# enteric_CH4=enteric_CH4,
enteric_CH4_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation_scenarios,
                   vars = "enteric_CH4",
                   method = "boxplot_density",
                   old_names = "enteric_CH4",
                   new_names = "CH4 Emissions from Enteric Fermentation per Farm") + 
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank())

@@

enteric_CH4_dist_plot + milk_yield_dist_plot + 
  enteric_CH4_CO2eq_dist_plot + mm_CH4_CO2eq_dist_plot + 
  mm_N2O_CO2eq_dist_plot + feed_CO2_dist_plot + 
  on_farm_dist_plot + 
plot_layout(guides = "collect", design = layout)

ggsave("Fig_1_GHG_simulation_outputs.png")
```

# THIS STILL NEEDS IMPROVEMENT - e.g. units are missing, and plot titles don't fit (use concise symbols with subscripts etc.)

## Possible errors resulting from input uncertainty

To illustrate the possible influence of input uncertainty on estimates of total farm emissions, we show the example of farm #5 from the household survey. This farm has 6 head of cattle, including 3 lactating cows. We simulate 1000 versions of this farm.

```{r, warning=FALSE}

scenarios<-read.csv("data/farm_scenarios.csv", fileEncoding = "UTF-8-BOM")
farm_5<-scenarios[,c(1,2,7)]

farm5_emissions<-scenario_mc(base_estimate = estimate_read_csv("data/livestock_ghg_input_table.csv"),
                             scenarios = farm_5,
                             model_function = ghg_emissions,
                             numberOfModelRuns = 1000,
                             functionSyntax = "plainNames")

plot_distributions(mcSimulation_object = farm5_emissions,
                   vars = "on_farm",
                   method = "boxplot_density",
                   old_names = "on_farm",
                   new_names = "Total Carbon Dioxide Equivalent Emissions from Livestock") + 
  theme(axis.title.y=element_blank())

ggsave("Fig_2_uncertainty_illustration.png")

# NEED TO ADD THE USUAL PRECISE VALUE HERE FOR COMPARISON (also need to add units and possible fix up other things)

This

```{r}
farm5_emissions$y[,"CO2eq_per_milk"]<-farm5_emissions$y$pc_on_farm/farm5_emissions$y$pc_milk_yield

ggplot(data=farm5_emissions$y, aes(CO2eq_per_milk)) +
geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") +
  ylab("Relative probability") +
    xlab(expression(Emission~density~of~milk~production~(XX~CO[2]-eq~kg^-1)))

ggsave("Fig_3_emission_density_farm5.png")

# UNIT NEEDS TO BE ADDED HERE (and checked for milk, though this is probably right)
# ALSO need to add the usual precise value estimate

# Population-wide emission density of milk production

Since the emission density of each household thus involves considerable uncertainty, such uncertainty should also be expected at the population level.

We can illustrate this uncertainty using a scatter plot. The following figure shows all combinations of total milk yield and total greenhouse gas emissions across all households.

```{r}
uncertainty::varscatter(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
```

# NICE PLOT, but would be nicer still if this was also a ggplot figure. Kinda inconsistent if it isn't. Also can't save it with ggsave, so will need another way.

@@

```{r}
uncertainty::varscatter(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
```

An indicator that is often of interest is the ratio between these two variables: total emissions divided by the milk yield. We'll calculate this next. We then have to remove all households that don't produce any milk, because this ratio would be infinity for them.

```{r}
GHG_simulation_scenarios$y[,"CO2eq_per_milk"]<-GHG_simulation_scenarios$y$pc_on_farm/GHG_simulation_scenarios$y$pc_milk_yield
GHG_simulation_scenarios$y<-GHG_simulation_scenarios$y[is.finite(GHG_simulation_scenarios$y$CO2eq_per_milk),]

Now we can plot the emissions intensity vs. the milk yield.

uncertainty::varscatter(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
                        out_var = GHG_simulation_scenarios$y$CO2eq_per_milk,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~intensity~of~milk~production~(kg~CO[2]-eq~kg^{-1})))
```

This can also be illustrated by a density surface.

```{r varkernel}
uncertainty::varkernel(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
                        out_var = GHG_simulation_scenarios$y$CO2eq_per_milk,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~intensity~of~milk~production~(kg~CO[2]-eq~kg^{-1})))
```

# THIS LOOKS BAD, BECAUSE OF Low-Milk-high-emissions households. Would be good to break the axis - any ideas?

The milk yield per capita is sometimes proposed as an indicator of greenhouse gas emissions. We can now examine the merits of this indicator. Essentially, its usefulness hinges on its relationship with greenhouse gas emission intensity. We can investigate this by taking slices through the density surface.

@@

```{r varkernelslicerange}
slice<-uncertainty::varslice_resample(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
                                 out_var = GHG_simulation_scenarios$y$CO2eq_per_milk, 
                                 expectedin_var=2000,
                           n = 1000,
                           n_samples = 1000,
                           out_var_sampling = (max(out_var)-min(out_var))/1000)

graphics::plot(x = slice$slice$Output_values, 
                 y = slice$slice$Relative_probability, 
                 type = "l", 
                 ylab = "Relative probability", 
                 xlab = "Emissions intensity", 
                 col = "seagreen", 
                 lwd = 2)

@@

# THIS WON'T REALLY WORK... peak is too high, everything else becomes invisible
